---
title: "A few beers later... `r emo::ji('clinking_beer_mugs')`"
params:
  data_file: ./data/allBeers.csv
  anonymize: 1
output:
  html_document:
    theme: paper
    highlight: kate
    toc: false
    includes:
      in_header: googleAnalytics.html
---

```{r setup, include=FALSE}
# Include and silence all chunks
knitr::opts_chunk$set(include = T, echo = F, warning = F, message = F, fig.width = 12)

# Load libraries
library(tidyverse)
library(ggthemes)  # colors
library(cowplot)  # themes
library(GGally)  # just for pairs plot (manually do?)
library(DT)  # javascript tables
library(emo)  # emojis (has to be installed with devtools)
library(randomNames)  # for making fake names

# Load data
all_beers <- data.table::fread(params$data_file) %>%
  dplyr::select(-V1) %>%
  dplyr::group_by(beer.beer_name, brewery.brewery_name, beer.bid) %>%
  dplyr::mutate(beer.rating_score = max(beer.rating_score)) %>%
  dplyr::ungroup()
n_users <- length(unique(all_beers$user))

# Anonymizes data if specified
if (as.numeric(params$anonymize) == 1) {
  
  # Maintains 3 original names for consistency otherwise makes random
  if (n_users == 3) {
    fake_users <- c("Alice", "Bob", "Carl") 
  } else {
    fake_users <- randomNames::randomNames(n = n_users, which.names = "first")
  }
  
  # Scrubs usernames from report
  all_beers <- all_beers %>%
    dplyr::right_join(dplyr::distinct(., user) %>%
                        dplyr::mutate(id = rank(user))) %>%
    dplyr::mutate(user = fake_users[id]) %>%
    dplyr::select(-id)
}

# List of users
users <- distinct(all_beers, user)$user

# Set plot theme to "cowplot"
theme_set(theme_cowplot())
```

# How do our distributions stack up? `r emo::ji('bar_chart')` {.tabset .tabset-pills}

## Distributions

```{r distributions}
# Violin plot with boxplot inside
all_beers %>%
  ggplot(aes(x = user, y = rating_score, fill = user)) +
  geom_violin() +
  geom_boxplot(width = 0.2, outlier.alpha = 0) +
  scale_fill_tableau() + ylim(c(0, 5)) +
  labs(x = "User", y = "Rating")
```

## Do we agree on ratings?

```{r correlations, fig.height=8}
# Pairs plot
all_beers %>%
  dplyr::select(user, beer.beer_name, brewery.brewery_name, rating_score, beer.rating_score) %>%
  dplyr::mutate(beer.rating_score = ifelse(beer.rating_score == 0, NA, beer.rating_score)) %>%
  tidyr::pivot_wider(names_from = "user", values_from = "rating_score") %>%  # make wide table for ggpairs
  dplyr::rename(global = beer.rating_score) %>%
  GGally::ggpairs(columns = c(3:ncol(.)))
```

## Controversial beers

```{r controversial}
# Find beers with largest difference between our ratings
controversial <- all_beers %>%
  dplyr::select(beer.beer_name, brewery.brewery_name, user, rating_score, beer.rating_score) %>%
  dplyr::group_by(beer.beer_name, brewery.brewery_name) %>%
  dplyr::mutate(max_diff = max(rating_score) - min(rating_score)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = "user", values_from = "rating_score") %>%
  dplyr::filter(max_diff >= 1)

# Make into interactive table
DT::datatable(controversial %>% dplyr::arrange(desc(max_diff)), style="bootstrap", width="100%",
              options = list(lengthChange = FALSE, scrollY = "300px", paging = FALSE),
              colnames = c("Beer", "Brewery", "Rating", "Maximum difference", colnames(controversial)[5:ncol(controversial)]),
              filter = "top") %>%
  DT::formatRound(columns = c("beer.rating_score", "max_diff"), digits = 3)
```

## Differences from global

```{r global diffs}
# Difference from global average (boxplot)
all_beers %>%
  dplyr::filter(beer.rating_score != 0) %>%
  dplyr::select(user, beer.beer_name, brewery.brewery_name, rating_score, beer.rating_score) %>%
  dplyr::mutate(glob_diff = rating_score - beer.rating_score) %>%  # calculate difference
  ggplot(aes(x = user, y = glob_diff)) +
  geom_boxplot(outlier.alpha = 0) + geom_jitter(aes(color = user)) +
  scale_color_tableau() +
  labs(x = "User", y = "Rating difference")
```

# Breweries `r emo::ji('factory')` {.tabset .tabset-pills}

Ratings by brewery for each user. User must have had 5 or more unique beers from the brewery.

## Top rated

```{r breweries, fig.height=12}
# Our ratings higher than 4.5
p1 <- all_beers %>%
  dplyr::filter(rating_score >= 4.5) %>%
  dplyr::group_by(brewery.brewery_name) %>%
  dplyr::mutate(n = n()) %>%
  ggplot(aes(y = reorder(brewery.brewery_name, n), fill = user)) +
  geom_bar() + scale_fill_tableau() +
  labs(x = "Number rated higher than 4.5", y = "Brewery", fill = "User")
# Global ratings higher than 4.25
p2 <- all_beers %>%
  dplyr::filter(beer.rating_score >= 4.25) %>%
  dplyr::distinct(beer.rating_score, brewery.brewery_name) %>%
  dplyr::group_by(brewery.brewery_name) %>%
  dplyr::mutate(n = n()) %>%
  ggplot(aes(y = reorder(brewery.brewery_name, n))) +
  geom_bar() + scale_fill_tableau() +
  labs(x = "Number with average global higher than 4.25", y = "Brewery")
cowplot::plot_grid(p2, p1, ncol = 1, rel_heights = c(6, 10))
```

## Average ratings {.tabset .tabset-dropdown}

```{r brewery averages}
# Ratings by brewery
brewery_ratings <- all_beers %>%
  dplyr::group_by(user, brewery.brewery_name) %>%
  dplyr::mutate(n = n()) %>%
  dplyr::ungroup() %>%
  dplyr::filter(n >= 5)  # user has had >= 5

# Loop through each  user and make plot
brewery_plots <- list()  # store plots
for (i in 1:length(users)) {
  p <- brewery_ratings %>%
    dplyr::filter(user == users[i]) %>%  # filter to current user
    ggplot(aes(x = rating_score, y = fct_reorder(brewery.brewery_name, rating_score, median),
               color = brewery.location.brewery_state)) +
    geom_boxplot(outlier.alpha = 0) + geom_jitter() +
    scale_color_tableau(palette = "Tableau 20") +
    xlim(c(0, 5)) +
    labs(x = "Rating", y = "Brewery", color = "State")
  brewery_plots[[i]] <- p
}
```

```{r brewery averages plot, results='asis', fig.height=8}
# Print all plots in their own subheading
for (i in 1:length(users)) {
  cat(sprintf('\n\n### %s {.tabset .tabset-pills}\n\n', users[i], '\n\n'))
  print(brewery_plots[[i]])
  cat("\n\n")
}
```

## Brewery rating table

```{r brewery table}
# Table of brewery ratings
brew_tab <- all_beers %>%
  dplyr::group_by(user, brewery.brewery_name) %>%
  dplyr::summarise(med_rating = median(rating_score), mean_rating = mean(rating_score), n = n(), .groups = "drop") %>%
  dplyr::filter(n >= 5) %>%  # user has  had >= 5
  dplyr::group_by(user) %>%
  dplyr::mutate(rank_med = rank(desc(med_rating), ties.method = "min"),
                rank_mean = rank(desc(mean_rating), ties.method = "min")) %>%  # rank breweries (ties get same value)
  dplyr::ungroup()

# Make into interactive table
DT::datatable(brew_tab %>% dplyr::arrange(rank_med), style="bootstrap", width="100%",
              options = list(lengthChange = FALSE, scrollY = "300px", paging = FALSE),
              colnames = c("User", "Brewery", "Median", "Mean", "Beers had", "Rank (median)", "Rank (mean)"),
              filter = "top") %>%
  DT::formatRound(columns = c("mean_rating", "med_rating"), digits = 3)  #round values
```

# Styles `r emo::ji('woman_dancing')` {.tabset .tabset-pills}

Ratings by style for each user. Users must have had 5 different beers of that style. **Note:** "meta" style refers to the  first part of the style while substyle refers to the entire style (e.g. _IPA - Imperial/Double_ would have a meta style of _IPA_ and a "substyle" style of _IPA - Imperial/Double_).

## Meta styles plots {.tabset .tabset-dropdown}

Styles broken down by first part of style annotation

```{r meta style}
# Styles by first word ("meta")
meta_style_tab <- all_beers %>%
  dplyr::mutate(meta_style = str_trim(word(beer.beer_style, 1, sep = fixed("-")))) %>%
  dplyr::group_by(user, meta_style) %>%
  dplyr::mutate(n = n()) %>%
  dplyr::ungroup() %>%
  dplyr::filter(n >= 5)

# Loop through users and make plots
style_plots <- list()
for (i in 1:length(users)) {
  p <- meta_style_tab %>%
    dplyr::filter(user == users[i]) %>%
    ggplot(aes(x = rating_score, y = fct_reorder(meta_style, rating_score, median))) +
    geom_boxplot(outlier.alpha = 0) +
    geom_jitter() +
    xlim(c(0, 5)) +
    labs(x = "Rating", y = "Style")
  style_plots[[i]] <- p
}
```

```{r meta style plot, results='asis', fig.height=8}
# Print each plot in own subheading
for (i in 1:length(users)) {
  cat(sprintf('\n\n### %s {.tabset .tabset-pills}\n\n', users[i], '\n\n'))
  print(style_plots[[i]])
  cat("\n\n")
}
```

## Meta styles table

```{r meta style table}
# Ratings by meta style
meta_table <- meta_style_tab %>%
  dplyr::group_by(user, meta_style) %>%
  dplyr::summarise(med_rating = median(rating_score), mean_rating = mean(rating_score), n = n(), .groups = "drop") %>%
  dplyr::group_by(user) %>%
  dplyr::mutate(rank_med = rank(desc(med_rating), ties.method = "min"),
                rank_mean = rank(desc(mean_rating), ties.method = "min")) %>%
  dplyr::ungroup()

# Make into interactive table
DT::datatable(meta_table %>% dplyr::arrange(rank_med), style="bootstrap", width="100%",
              options = list(lengthChange = FALSE, scrollY = "300px", paging = FALSE),
              colnames = c("User", "Style", "Median", "Mean", "Beers had", "Rank (median)", "Rank (mean)"),
              filter = "top") %>%
  DT::formatRound(columns = c("mean_rating", "med_rating"), digits = 3)
```

## Substyles table

More detailed style information

```{r granular style table}
# Ratings by granular style
style_tab <- all_beers %>%
  dplyr::group_by(user, beer.beer_style) %>%
  dplyr::summarise(med_rating = median(rating_score), mean_rating = mean(rating_score), n = n(), .groups = "drop") %>%
  dplyr::filter(n >= 5) %>%
  dplyr::group_by(user) %>%
  dplyr::mutate(rank_med = rank(desc(med_rating), ties.method = "min"),
                rank_mean = rank(desc(mean_rating), ties.method = "min")) %>%
  dplyr::ungroup()

# Make into interactive table
DT::datatable(style_tab %>% dplyr::arrange(rank_med), style="bootstrap", width="100%",
              options = list(lengthChange = FALSE, scrollY = "300px", paging = FALSE),
              colnames = c("User", "Substyle", "Median", "Mean", "Beers had", "Rank (median)", "Rank (mean)"),
              filter = "top") %>%
  DT::formatRound(columns = c("mean_rating", "med_rating"), digits = 3)
```

# ABV and IBU `r emo::ji('nauseated_face')` {.tabset .tabset-pills}

```{r abv ibu summary, fig.height=10}
# Pairs plot relating our ratings to ABV and IBU
all_beers %>%
  dplyr::select(user, beer.beer_name, brewery.brewery_name, rating_score,
                beer.beer_abv, beer.beer_ibu, beer.rating_score) %>%
  dplyr::mutate(beer.rating_score = ifelse(beer.rating_score == 0, NA, beer.rating_score),
                beer.beer_ibu = ifelse(beer.beer_ibu == 0, NA, beer.beer_ibu),
                beer.beer_abv = ifelse(beer.beer_abv == 0, NA, beer.beer_abv)) %>%
  tidyr::pivot_wider(names_from = "user", values_from = "rating_score") %>%
  dplyr::rename(global = beer.rating_score, ABV = beer.beer_abv, IBU = beer.beer_ibu) %>%
  GGally::ggpairs(columns = c(3:ncol(.)))
```

# Prediction `r emo::ji('monocle')`

I fit a 10-fold cross-validated random forest to predict each user's ratings. Reported here are the Pearson score of each model as well as estimated importance of each feature in prediction (and the associated standard deviation).

```{r prediction}
predictions <- list()  # store actual predictions
results <- list()  # store model stats

# Fit a 10-fold CV random forest for each user
for (i in 1:length(users)) {
  
  # Filter to user and relevant columns:
  # beer id, rating, global, num ratings abv, ibu, style, brewery type, brewery state
  user <- all_beers %>%
    dplyr::filter(user == users[[i]]) %>%
    dplyr::select(beer.bid, rating_score, beer.beer_abv, beer.beer_ibu,
                  beer.beer_style, beer.rating_score, brewery.brewery_type,
                  brewery.location.brewery_state, beer.rating_count)
  
  # Make CV folds
  folds <- caret::createFolds(user$rating_score, k = 10)
  
  preds <- user$rating_score
  importances <- list()
  
  # Loop through folds for CV
  for (k in 1:length(folds)) {
    
    # Test and training sets
    fold <- folds[[k]]
    test <- user[fold, ]
    train <- dplyr::anti_join(user, test)
    
    # Fit RF
    res <- ranger::ranger(rating_score ~ ., data = train %>%
                            dplyr::select(-beer.bid),
                          importance = "impurity", respect.unordered.factors = T)
    
    # Store model statistics and predictions
    preds[fold] <- predict(res, test %>% dplyr::select(-rating_score, -beer.bid))$predictions
    imps <- tibble::tibble(feature = names(res$variable.importance),
                           RF.imp = res$variable.importance / sum(res$variable.importance),
                           fold = k)
    importances[[k]] <- imps
  }
  
  # Combine across folds
  importances <- dplyr::bind_rows(importances) %>%
    dplyr::distinct(feature, RF.imp, fold) %>%
    reshape2::acast(feature ~ fold, value.var = "RF.imp")
  
  # Feature importances
  imp_table <- tibble::tibble(feature = rownames(importances),
                              imp.mean = importances %>%
                                apply(1, function(x) mean(x, na.rm = T)),
                              imp.sd = importances %>%
                                apply(1, function(x) sd(x, na.rm = T)),
                              imp.stability = importances %>%
                                apply(1, function(x) mean(!is.na(x)))) %>%
    dplyr::filter(feature != "(Intercept)") %>%
    dplyr::arrange(desc(imp.mean)) %>%
    dplyr::mutate(rank = 1:n()) %>%
    dplyr::mutate(pearson = cor(preds, user$rating_score), user = users[[i]])
  
  # Output to final list
  results[[i]] <- imp_table
  predictions[[i]] <- tibble(predicted = preds, actual = user$rating_score,
                             bid = user$beer.bid, user = users[[i]])
}

# Combine across users
results <- dplyr::bind_rows(results); predictions <- dplyr::bind_rows(predictions)

# Pearson scores plot
p1 <- results %>%
  dplyr::distinct(user, pearson) %>%
  ggplot(aes(x = user, y = pearson, fill = user)) +
  geom_bar(stat = "identity") +
  scale_fill_tableau() + theme(legend.position = "none") +
  labs(x = "User", y = "Pearson score")

# Feature importance plot
p2 <- results %>%
  dplyr::distinct(user, feature, imp.mean, rank) %>%
  ggplot(aes(x = imp.mean, y = fct_reorder(feature, rank, "median", .desc = T), fill = user)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_tableau() +
  labs(y = "Feature", x = "Mean feature importance", fill = "User")
cowplot::plot_grid(p1, p2, rel_widths = c(1, 3))
```

```{r prediction table}
# Table of random forest results
DT::datatable(results %>%
                dplyr::arrange(rank) %>%
                dplyr::select(c(7, 6, 1, 2, 3)),
              style="bootstrap", width="100%",
              options = list(lengthChange = FALSE, scrollY = "300px", paging = FALSE),
              colnames = c("User", "Pearson Score", "Feature", "Importance", "Importance deviation"),
              filter = "top") %>%
  DT::formatRound(columns = c("pearson", "imp.mean", "imp.sd"), digits = 3)
```
